---
title: "Introduction to clustRcompaR"
author: "Joshua Rosenberg and Alex Lishinski"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to clustRcompaR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# clustRcompaR

An `R` package to cluster and compare text data.

## Background 
[Document clustering](https://en.wikipedia.org/wiki/Document_clustering) is a common technique to discover topics in a corpus of texts. This package uses functions from the [`quanteda`](https://github.com/kbenoit/quanteda) `R` package as the basis for two functions, `cluster()` and `compare(), to make document clustering and comparing topics identified through document clustering across factors straightforward.

## Installation

Because this package is in development and is not yet available on CRAN, to install it, first install the `devtools` package using `install.packages("devtools")`, followed by the function `devtools::install_github("alishinski/clustRcompaR")`. After installing the package, use `library(clustRcompaR)` to load it in each session.

## Workflow

* First, use `cluster()` on a `data.frame` with the first column a `vector` of `strings` and any subsequent columns `vectors` of `factors`.
 
 * Optional arguments to the `cluster()` function include parameters for the minimum frequency with which a term must occur to be included in the analysis, the minimum number of terms in each document after processing, and additional stopwords. The output from the `cluster()` function can then be inspected to determine the interpretability of clusters and the suitability of the clustering solution.

* Next, use `compare()` with the output from the `cluster()` function along with a `string` for the factor to compare the frequency of clusters to. 
 
 * This output can be examined and used on its own, or be passed to two functions, `compare_plot()`, which plots the table using [`ggplot2`](https://github.com/hadley/ggplot2), or `compare_test()`, which performs a chi-square test of proportions on the table and indicates across which levels of the factor clusters appear more or less likely than expected.
 
## Example
 
Here is an example using the built-in `inaugural_addresses` dataset (from the quanteda package). This dataset consists of the inaugural addresses by every United States President.

First, we use `cluster()` to cluster the documents into three clusters. We include a new variable, `year_before_1900`, which we will later use to compare frequencies across clusters. Then we use `extract_terms()` to view the terms and term frequencies in the two clusters.

```{r example}
library(clustRcompaR)
library(dplyr)
library(quanteda)

d <- inaugural_addresses
d <- mutate(d, century = ifelse(Year < 1800, "17th",
                                ifelse(Year >= 1800 & Year < 1900, "18th",
                                       ifelse(Year >= 1900 & Year < 2000, "19th", "20th"))))

three_clusters <- cluster(d, century, n_clusters = 3)

extract_terms(three_clusters)
```

Second, we use the `compare()` function to compare the frequency of clusters across a factor, in this case, the century. We can then use the `compare_plot()` or `compare_test()` (which uses a Chi-Square test) function. 

```{r, example-p2}
three_clusters_comparison <- compare(three_clusters, "century")

compare_plot(three_clusters_comparison)
```
